### 2015-07-21 Appendix/Appendix1.Rnw
appendix A1 edits to make sure no single lines at beginning or end of pages

i. Two different samples will be taken from the same population of test scores where the population mean and standard deviation are unknown. The first sample will have 25 data values, and the second sample will have 64 data values. A 95\% confidence interval will be constructed for each sample to estimate the population mean. Which confidence interval would you expect to have greater precision (a smaller width) for estimating the population mean?

a. I expect the confidence interval based on the sample of 64 data values to be more precise.\\
b. I expect both confidence intervals to have the same precision.\\
c. I expect the confidence interval based on the sample of 25 data values to be more precise.\\

ii. A 95\% confidence interval is computed to estimate the mean household income for a city. Which of the following values will definitely be within the limits of this confidence interval?

a. The population mean\\
b. The sample mean\\
c. The standard deviation of the sample mean\\
d. None of the above\\

iii. Each of the 110 students in a statistics class selects a different random sample of 35 Quiz scores from a population of 5000 scores they are given. Using their data, each student constructs a 90\% confidence interval for $\mu$ the average Quiz score of the
5000 students. Which of the following conclusions is correct?

a. About 10\% of the sample means will not be included in the confidence intervals.\\
b. About 90\% of the confidence intervals will contain $\mu$.\\
c. It is probable that 90\% of the confidence intervals will be identical.\\
d. About 10\% of the raw scores in the samples will not be found in these confidence intervals\\

iv. A 95\% confidence interval for the mean reading achievement score for a population of third grade students is (43, 49). The margin of error of this interval is:

a. 5\\
b. 3\\
c. 6\\

v. Justin and Hayley conducted a mission to a new planet, Planet X, to study arm length. They took a random sample of 100 Planet X residents and calculated a 95\% confidence interval for the mean arm length. What does a 95\% confidence interval for arm length tell us in this case? Select the best answer:

a. I am 95\% confident that this interval includes the sample mean (x) arm length.\\
b. I am confident that most (95\%) of all Planet X residents will have an arm length within this interval.\\
c. I am 95\% confident that most Planet X residents will have arm lengths within this interval.\\
d. I am 95\% confident that this interval includes the population mean arm length.\\

vi. Suppose that a random sample of 41 state college students is asked to measure the length of their right foot in centimeters. A 95\% confidence interval for the mean foot length for students at this university turns out to be (21.709, 25.091). If instead a 90\% confidence interval was calculated, how would it differ from the 95\% confidence interval?

a. The 90\% confidence interval would be narrower.\\
b. The 90\% confidence interval would be wider.\\
c. The 90\% confidence interval would be the same as the 95\% confidence interval.\\

vii. A pollster took a random sample of 100 students from a large university and computed a confidence interval to estimate the percentage of students who were planning to vote in the upcoming election. The pollster felt that the confidence interval was too wide to provide a precise estimate of the population parameter. What could the pollster have done to produce a narrower confidence interval that would produce a more precise estimate of the percentage of all university students who plan to vote in the upcoming election?

a. Increase the sample size to 150.\\
b. Increase the confidence level to 99\%.\\
c. Both a and b\\
d. None of the above\\

viii. A newspaper article states with 95\% confidence that 55\% to 65\% of all high school students in the United States claim that they could get a hand gun if they wanted one. This confidence interval is based on a poll of 2000 high school students in Detroit. How would you interpret the confidence interval from this newspaper article?

a. 95\% of large urban cities in the United States have 55\% to 65\% high school students who could get a hand gun.\\
b. If we took many samples of high school students from different urban cities, 95\% of the samples would have between 55\% and 65\% high school students who could get hand guns.\\
c. You cannot use this confidence interval to generalize to all teenagers in the United States because of the way the sample was taken.\\
d. We can be 95\% confident that between 55\% and 65\% of all United States high school students could get a hand gun.\\

ix. The Gallup poll (August 23, 2002) reported that 53\% of Americans said they would favor sending American ground troops to the Persian Gulf area in an attempt to remove Hussein from power. The poll also reported that the margin of error for this poll was 4\%. What does the margin of error of 4\% indicate?

a. There is a 4\% chance that the estimate of 53\% is wrong.\\
b. The percent of Americans who are in favor is probably higher than 53\% and closer to 57\%.\\
c. The percent of Americans who are in favor is estimated to be between 49\% and 57\%.\\

x. Suppose two researchers want to estimate the proportion of American college students who favor abolishing the penny. They both want to have about the same margin of error to estimate this proportion. However, Researcher 1 wants to estimate with 99\% confidence and Researcher 2 wants to estimate with 95\% confidence. Which researcher would need more students for her study in order to obtain the desired margin of error?

a. Researcher 1.\\
b. Researcher 2.\\
c. Both researchers would need the same number of subjects.\\
d. It is impossible to obtain the same margin of error with the two different confidence levels.\\

@@

i. The makers of Mini-Oats cereal have an automated packaging machine that is set to fill boxes with 24 ounces of cereal. At various times in the packaging process, a random sample of 100 boxes is taken to see if the machine is filling the boxes with an average of 24 ounces of cereal. Which of the following is a statement of the null hypothesis being tested?

a. The machine is filling the boxes with the proper amount of cereal.\\
b. The machine is not filling the boxes with the proper amount of cereal.\\
c. The machine is not putting enough cereal in the boxes.\\

ii. A research article gives a p-value of .001 in the analysis section. Which definition of a p-value is the most accurate?

a. the probability that the observed outcome will occur again.\\
b. the probability of observing an outcome as extreme or more extreme than the one observed if the null hypothesis is true.\\
c. the value that an observed outcome must reach in order to be considered significant under the null hypothesis.\\
d. the probability that the null hypothesis is true.\\

iii. If a researcher was hoping to show that the results of an experiment were statistically significant they would prefer:

a. a large p-value\\
b. a small p-value\\
c. p-values are not related to statistical significance\\

iv. A researcher compares men and women on 100 different variables using a difference in means t-test.  He sets the level of significance at 0.05 and then carries out 100 independent t-tests (one for each variable) on these data.  If, for each test, the null hypothesis is actually true, about how many ``statistically significant'' results will be produced?

a. 0\\
b. 5\\
c. 10\\
d. none of the above\\

Problems (v) and (vi) refer to the following situation: Food inspectors inspect samples of food products to see if they are safe. This can be thought of as a hypothesis test where Null: the food is safe, and Alternative: the food is not safe. Identify each of the following statements as a Type I or a Type II error.

v. The inspector says the food is safe but it actually is not safe.

a. Type I\\
b. Type II\\

vi. The inspector says the food is not safe but is actually safe.

a. Type I\\
b. Type II\\

vii. A newspaper article claims that the average age for people who receive food stamps is 40 years. You believe that the average age is less than that. You take a random sample of 100 people who receive food stamps, and find their average age to be 39.2 years. You find that this is significantly lower than the age of 40 stated in the article (p-value < .05). What would be an appropriate interpretation of this result?

a. The statistically significant result indicates that the majority of people who receive food stamps is younger than 40.\\
b. Although the result is statistically significant, the difference in age is not of practical importance.\\
c. An error must have been made. This difference is too small to be statistically significant.\\

viii. A newspaper article stated that the US Supreme Court received 812 letters from around the country on the subject of whether to ban cameras from the courtroom. Of these 812 letters, 800 expressed the opinion that cameras should be banned. A statistics student was going to use this sample information to conduct a test of significance of whether more than 95\% of all American adults feel that cameras should be banned from the courtroom. What would you tell this student?

a. This is a large enough sample to provide an accurate estimate of the American public's opinion on the issue.\\
b. The necessary conditions for a test of significance are not satisfied, so no statistical test should be performed.\\
c. With such a large number of people favoring the notion that cameras be banned, there is no need for a statistical test.\\

ix. A researcher conducts an experiment on human memory and recruits 15 people to participate in her study. She performs the experiment and analyzes the results. She obtains a p-value of .17. Which of the following is a reasonable interpretation of her results?

a. This proves that her experimental treatment has no effect on memory.\\
b. There could be a treatment effect, but the sample size was too small to detect it.\\
c. She should reject the null hypothesis.\\
d. There is evidence of a small effect on memory by her experimental treatment.\\

x. It is reported that scores on a particular test of historical trivia given to high school students are approximately normally distributed with a mean of 85. Mrs. Rose believes that her 5 classes of high school seniors will score significantly better than the national average on this test. At the end of the semester, Mrs. Rose administers the historical trivia test to her students. The students score an average of 89 on this test. After conducting the appropriate statistical test, Mrs. Rose finds that the p-value is .0025. Which of the following is the best interpretation of the p-value?

a. A p-value of .0025 provides strong evidence that Mrs. Rose's class outperformed high school students across the nation.\\
b. A p-value of .0025 indicates that there is a very small chance that Mrs. Rose's class outperformed high school students across the nation.\\
c. A p-value of .0025 provides evidence that Mrs. Rose is an exceptional teacher who was able to prepare her students well for this national test.\\
d. None of the above.\\

@@

\textbf{1. [28 points] Linear Regression:} A study was conducted in England in 1989 investigating the relationship between smoking and lung cancer mortality within different occupational groups of males. The data include a smoking index and a lung cancer mortality index for men in \Sexpr{nrow(smokedie)} occupational groups in England. \\
\indent The smoking index measures the percentage of the number of cigarettes smoked per day by men in the particular occupational group compared to the average number of cigarettes smoked per day by all men. The units associated with the smoking index are percent of average cigarettes smoked (\% avg CS).  The mortality index measures the percentage of the rate of deaths from lung cancer among men in the particular occupational group compared to the rate of deaths from lung cancer among all men. The units associated with the mortality index are percent of average lung cancer deaths (\% avg LCD).  \\
\indent The data for the \Sexpr{nrow(smokedie)} occupations are displayed in the scatterplot below. Use the information from the scatterplot, the sample statistics and linear regression equation listed below to answer all parts of this problem. \\
<< 'regprobplot' , echo=FALSE, warning=FALSE, fig.width=6, fig.height=6, out.width='.5\\linewidth', fig.pos='H',fig.align='center'>>=
mod1 <- lm(Mortality ~ Smoking, smokedie)
#str(summary(mod1))
int <- mod1$coefficients[1]
slo <- mod1$coefficients[2]

>>>>

i. Two different samples will be taken from the same population of test scores where the population mean and standard deviation are unknown. The first sample will have 25 data values, and the second sample will have 64 data values. A 95\% confidence interval will be constructed for each sample to estimate the population mean. Which confidence interval would you expect to have greater precision (a smaller width) for estimating the population mean?\\
a. I expect the confidence interval based on the sample of 64 data values to be more precise.\\
b. I expect both confidence intervals to have the same precision.\\
c. I expect the confidence interval based on the sample of 25 data values to be more precise.\\

ii. A 95\% confidence interval is computed to estimate the mean household income for a city. Which of the following values will definitely be within the limits of this confidence interval?\\
a. The population mean\\
b. The sample mean\\
\newpage
\noindent c. The standard deviation of the sample mean\\
d. None of the above\\

iii. Each of the 110 students in a statistics class selects a different random sample of 35 Quiz scores from a population of 5000 scores they are given. Using their data, each student constructs a 90\% confidence interval for $\mu$ the average Quiz score of the
5000 students. Which of the following conclusions is correct?\\
a. About 10\% of the sample means will not be included in the confidence intervals.\\
b. About 90\% of the confidence intervals will contain $\mu$.\\
c. It is probable that 90\% of the confidence intervals will be identical.\\
d. About 10\% of the raw scores in the samples will not be found in these confidence intervals\\

iv. A 95\% confidence interval for the mean reading achievement score for a population of third grade students is (43, 49). The margin of error of this interval is:\\
a. 5\\
b. 3\\
c. 6\\

v. Justin and Hayley conducted a mission to a new planet, Planet X, to study arm length. They took a random sample of 100 Planet X residents and calculated a 95\% confidence interval for the mean arm length. What does a 95\% confidence interval for arm length tell us in this case? Select the best answer:\\
a. I am 95\% confident that this interval includes the sample mean (x) arm length.\\
b. I am confident that most (95\%) of all Planet X residents will have an arm length within this interval.\\
c. I am 95\% confident that most Planet X residents will have arm lengths within this interval.\\
d. I am 95\% confident that this interval includes the population mean arm length.\\

vi. Suppose that a random sample of 41 state college students is asked to measure the length of their right foot in centimeters. A 95\% confidence interval for the mean foot length for students at this university turns out to be (21.709, 25.091). If instead a 90\% confidence interval was calculated, how would it differ from the 95\% confidence interval?\\
a. The 90\% confidence interval would be narrower.\\
b. The 90\% confidence interval would be wider.\\
c. The 90\% confidence interval would be the same as the 95\% confidence interval.\\

vii. A pollster took a random sample of 100 students from a large university and computed a confidence interval to estimate the percentage of students who were planning to vote in the upcoming election. The pollster felt that the confidence interval was too wide to provide a precise estimate of the population parameter. What could the pollster have done to produce a narrower confidence interval that would produce a more precise estimate of the percentage of all university students who plan to vote in the upcoming election?\\
a. Increase the sample size to 150.\\
b. Increase the confidence level to 99\%.\\
c. Both a and b\\
d. None of the above\\

viii. A newspaper article states with 95\% confidence that 55\% to 65\% of all high school students in the United States claim that they could get a hand gun if they wanted one. This confidence interval is based on a poll of 2000 high school students in Detroit. How would you interpret the confidence interval from this newspaper article?\\
a. 95\% of large urban cities in the United States have 55\% to 65\% high school students who could get a hand gun.\\
b. If we took many samples of high school students from different urban cities, 95\% of the samples would have between 55\% and 65\% high school students who could get hand guns.\\
c. You cannot use this confidence interval to generalize to all teenagers in the United States because of the way the sample was taken.\\
d. We can be 95\% confident that between 55\% and 65\% of all United States high school students could get a hand gun.\\

ix. The Gallup poll (August 23, 2002) reported that 53\% of Americans said they would favor sending American ground troops to the Persian Gulf area in an attempt to remove Hussein from power. The poll also reported that the margin of error for this poll was 4\%. What does the margin of error of 4\% indicate?\\
a. There is a 4\% chance that the estimate of 53\% is wrong.\\
b. The percent of Americans who are in favor is probably higher than 53\% and closer to 57\%.\\
c. The percent of Americans who are in favor is estimated to be between 49\% and 57\%.\\

x. Suppose two researchers want to estimate the proportion of American college students who favor abolishing the penny. They both want to have about the same margin of error to estimate this proportion. However, Researcher 1 wants to estimate with 99\% confidence and Researcher 2 wants to estimate with 95\% confidence. Which researcher would need more students for her study in order to obtain the desired margin of error?\\
a. Researcher 1.\\
b. Researcher 2.\\
c. Both researchers would need the same number of subjects.\\
d. It is impossible to obtain the same margin of error with the two different confidence levels.\\

@@

i. The makers of Mini-Oats cereal have an automated packaging machine that is set to fill boxes with 24 ounces of cereal. At various times in the packaging process, a random sample of 100 boxes is taken to see if the machine is filling the boxes with an average of 24 ounces of cereal. Which of the following is a statement of the null hypothesis being tested?\\
a. The machine is filling the boxes with the proper amount of cereal.\\
b. The machine is not filling the boxes with the proper amount of cereal.\\
c. The machine is not putting enough cereal in the boxes.\\

ii. A research article gives a p-value of .001 in the analysis section. Which definition of a p-value is the most accurate?\\
a. the probability that the observed outcome will occur again.\\
b. the probability of observing an outcome as extreme or more extreme than the one observed if the null hypothesis is true.\\
c. the value that an observed outcome must reach in order to be considered significant under the null hypothesis.\\
d. the probability that the null hypothesis is true.\\

iii. If a researcher was hoping to show that the results of an experiment were statistically significant they would prefer:\\
a. a large p-value\\
b. a small p-value\\
c. p-values are not related to statistical significance\\

iv. A researcher compares men and women on 100 different variables using a difference in means t-test.  He sets the level of significance at 0.05 and then carries out 100 independent t-tests (one for each variable) on these data.  If, for each test, the null hypothesis is actually true, about how many ``statistically significant'' results will be produced?\\
a. 0\\
b. 5\\
c. 10\\
d. none of the above\\

Problems (v) and (vi) refer to the following situation: Food inspectors inspect samples of food products to see if they are safe. This can be thought of as a hypothesis test where Null: the food is safe, and Alternative: the food is not safe. Identify each of the following statements as a Type I or a Type II error.\\

v. The inspector says the food is safe but it actually is not safe.\\
a. Type I\\
b. Type II\\

vi. The inspector says the food is not safe but is actually safe.\\
a. Type I\\
b. Type II\\

vii. A newspaper article claims that the average age for people who receive food stamps is 40 years. You believe that the average age is less than that. You take a random sample of 100 people who receive food stamps, and find their average age to be 39.2 years. You find that this is significantly lower than the age of 40 stated in the article (p-value < .05). What would be an appropriate interpretation of this result?\\
a. The statistically significant result indicates that the majority of people who receive food stamps is younger than 40.\\
b. Although the result is statistically significant, the difference in age is not of practical importance.\\
c. An error must have been made. This difference is too small to be statistically significant.\\

viii. A newspaper article stated that the US Supreme Court received 812 letters from around the country on the subject of whether to ban cameras from the courtroom. Of these 812 letters, 800 expressed the opinion that cameras should be banned. A statistics student was going to use this sample information to conduct a test of significance of whether more than 95\% of all American adults feel that cameras should be banned from the courtroom. What would you tell this student?\\
a. This is a large enough sample to provide an accurate estimate of the American public's opinion on the issue.\\
b. The necessary conditions for a test of significance are not satisfied, so no statistical test should be performed.\\
c. With such a large number of people favoring the notion that cameras be banned, there is no need for a statistical test.\\

ix. A researcher conducts an experiment on human memory and recruits 15 people to participate in her study. She performs the experiment and analyzes the results. She obtains a p-value of .17. Which of the following is a reasonable interpretation of her results?\\
a. This proves that her experimental treatment has no effect on memory.\\
b. There could be a treatment effect, but the sample size was too small to detect it.\\
c. She should reject the null hypothesis.\\
d. There is evidence of a small effect on memory by her experimental treatment.\\

x. It is reported that scores on a particular test of historical trivia given to high school students are approximately normally distributed with a mean of 85. Mrs. Rose believes that her 5 classes of high school seniors will score significantly better than the national average on this test. At the end of the semester, Mrs. Rose administers the historical trivia test to her students. The students score an average of 89 on this test. After conducting the appropriate statistical test, Mrs. Rose finds that the p-value is .0025. Which of the following is the best interpretation of the p-value?\\
a. A p-value of .0025 provides strong evidence that Mrs. Rose's class outperformed high school students across the nation.\\
b. A p-value of .0025 indicates that there is a very small chance that Mrs. Rose's class outperformed high school students across the nation.\\
\newpage
\noindent c. A p-value of .0025 provides evidence that Mrs. Rose is an exceptional teacher who was able to prepare her students well for this national test.\\
d. None of the above.\\

@@

\textbf{1. [28 points] Linear Regression:} A study was conducted in England in 1989 investigating the relationship between smoking and lung cancer mortality within different occupational groups of males. The data include a smoking index and a lung cancer mortality index for men in \Sexpr{nrow(smokedie)} occupational groups in England. \\
\indent The smoking index measures the percentage of the number of cigarettes smoked per day by men in the particular occupational group compared to the average number of cigarettes smoked per day by all men. The units associated with the smoking index are percent of average cigarettes smoked (\% avg CS).  The mortality index measures the percentage of the rate of deaths from lung cancer among men in the particular occupational group compared to the rate of deaths from lung cancer among all men. The units associated with the mortality index are percent of average lung cancer deaths (\% avg LCD).  \\
\newpage
\indent The data for the \Sexpr{nrow(smokedie)} occupations are displayed in the scatterplot below. Use the information from the scatterplot, the sample statistics and linear regression equation listed below to answer all parts of this problem. \\
<< 'regprobplot' , echo=FALSE, warning=FALSE, fig.width=6, fig.height=6, out.width='.5\\linewidth', fig.pos='H',fig.align='center'>>=
mod1 <- lm(Mortality ~ Smoking, smokedie)
#str(summary(mod1))
int <- mod1$coefficients[1]
slo <- mod1$coefficients[2]
